MCC Preference Kids

### 2019-02-08 stats/kids/kids_Ex3_pref.Rmd
update kids model preictions

ggplot(data = p3, aes(x = age_num, y = correct_inf)) +
  geom_jitter(width = .025,height = .025, alpha = .2)+
  geom_smooth(method = "loess",  se = T,  alpha = .5, size = .4, span = 2)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion more informative")+
  theme_few() +
  facet_grid(speaker~alignment)+
  ylim(-0.05,1.05)+
  guides(alpha = F)

>>>>

ggplot(data = p3, aes(x = age_num, y = correct_inf)) +
  geom_jitter(width = .025,height = .025, alpha = .2)+
    geom_smooth(col = "black", size = 0.8, method = "glm", method.args = list(family = "binomial"), se = T,  alpha = .7)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion more informative")+
  theme_few() +
  facet_grid(speaker~alignment)+
  ylim(-0.05,1.05)+
  guides(alpha = F)

### 2019-02-08 stats/kids/kids_model_predictions.Rmd
update kids model preictions

```{r pretest data, include=FALSE}
pref_data <- read_csv(file="../data/kids_pref_data.csv")%>%
  filter(age != "2",
         trial_type != "train")%>%
  mutate(age_bin = factor(age),
         experiment = "preference_pretest",
         condition = ifelse(change == F, "same_speaker", "different_speaker"),
         agent = ifelse( change == F, agent, altAgent),
         minage = min(age_num))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct,minage)

@@

pre_data <- bind_rows(pref_data,
                      inf_data)
```

@@

@@

@@

@@

    return extend(
    // diff_speaker_params, same_speaker_params, 
      {
      diff_priorReg,
      same_priorReg,
      speakerOptimality,
      rsaPredictions_diff_cong,
      rsaPredictions_diff_incong,
      rsaPredictions_same_cong,
      rsaPredictions_same_incong,
      age_bin,
      so_slope,
      so_int,

model_pred <- readRDS("../saves/kids_model_predictions.rds")

# model_pred<- webppl(
#   program_code = paste(rsaUtils, rsaModel, modelPredWebppl , sep='\n'),
#   data = pre_data,
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 900, burn = 200, verbose = T)
# )

kids_model_pred <- model_pred%>%
  select(value) %>%
  map_df(bind_rows)%>%
  mutate(iteration = rep(1:3600,each = 40))%>%
  select(-diff_priorReg,-same_priorReg,-speakerOptimality,-so_slope,-so_int)%>%
  gather(condition, prop_informative, -iteration,-age_bin)%>%
  separate(condition, into = c("model", "Speaker", "Alignment"), sep="_")%>%
  mutate(Speaker = ifelse(Speaker == "diff", "different_speaker","same_speaker"),
         Alignment = ifelse(Alignment == "cong", "congruent","incongruent"))

kids_model_pred_so <- model_pred%>%
  select(value) %>%
  map_df(bind_rows)%>%
  mutate(iteration = rep(1:3600,each = 40))%>%
  select(so_slope,so_int,age_bin,iteration,speakerOptimality)

kids_model_pred_prior <- model_pred%>%
  select(value) %>%
  map_df(bind_rows)%>%
  mutate(iteration = rep(1:3600,each = 40))%>%
  select(age_bin,iteration,diff_priorReg,same_priorReg)%>%
  gather(condition, prop_preferred, -iteration,-age_bin)%>%
  mutate(condition = ifelse(condition == "diff_priorReg", "different_speaker","same_speaker"))

@@

    return extend(
    // diff_speaker_params, same_speaker_params, 
      {
      diff_priorReg,
      same_priorReg,
      speakerOptimality,
      noisyRsaPredictions_diff_cong,
      noisyRsaPredictions_diff_incong,
      noisyRsaPredictions_same_cong,
      noisyRsaPredictions_same_incong,
      age_bin,
      so_slope,
      so_int,
      noise
    })

@@

@@

```{r noisy combined model predictions}

```
# Sanity checks

@@

## importing data
d <- read_csv(file="../data/kids_ex3_data.csv")

>>>>

```{r pretest data, include=FALSE}

pref_data <- read_csv(file="../data/kids_pref_data.csv")%>%
  filter(age != "2",
         trial_type != "train")%>%
  mutate(age_bin = factor(age),
         experiment = "preference_pretest",
         condition = ifelse(change == F | change == "false", "same_speaker", "different_speaker"),
         agent = ifelse( change == F | change == "false", agent, altAgent),
         minage = min(age_num))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct,minage)

@@

pre_data <- bind_rows(pref_data,
                      inf_data)

ex3_data <- read_csv(file="../data/kids_ex3_data.csv")%>%
  filter(trial != "filler1",trial != "filler2")

@@

## Model parameters by age bin

```{r}
paramModelWebppl <- '

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}
var prefData = _.filter(dataFromR, {experiment: "preference_pretest"})
var infData = _.filter(dataFromR, {experiment: "informativeness_pretest"})
// make finer grained as needed
var binned_age_range = _.range(0, 2, 0.05)
var all_conditions = levels(prefData, "condition")
var foreach = function(fn, lst) {
    var foreach_ = function(i) {
        if (i < lst.length) {
            fn(lst[i]);
            foreach_(i + 1);
        }
    };
    foreach_(0);
};
var logistic = function(x) {1 / (1 + Math.exp(-x))}
var model  = function(){
  var so_slope = uniformDrift({a: -2, b: 2, width: 0.4})
  var so_int = uniformDrift({a: -2, b: 2, width: 0.4})
  foreach(function(row){
    var age = row.age_num
    var speakerOptimality = so_int  + so_slope * (age - infData[0].minage)

    var inf_priorProbs = [.5, .5, .5] 
    var rsaPredictions = pragmaticListener({label: "dax", point: 2 },
    inf_priorProbs, speakerOptimality)

    observe(rsaPredictions, row.correct)
  }, infData)
  var pref_params = map(function(cndtn){
    var conditionData = _.filter(prefData, {condition: cndtn})

    foreach(function(row){
      var age = row.age_num
      var priorReg = logistic(pref_int + pref_slope * (age - row.minage))
      var priorProbs= [1-priorReg, priorReg]
      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: prior_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      observe(priorPredictions, row.correct)
    }, conditionData)

  var diff_speaker_params = _.filter(pref_params, 
        {condition: "different_speaker"})[0]
  var same_speaker_params = _.filter(pref_params, 
        {condition: "same_speaker"})[0]

    return extend(
      {
      diff_priorReg,
      same_priorReg,
      speakerOptimality,
      age_bin,

```{r}
params_by_age<- readRDS("../saves/params_by_age.rds")

# params_by_age<- webppl(
#   program_code = paste(rsaUtils, rsaModel, paramModelWebppl , sep='\n'),
#   data = pre_data,
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
# )
# 
# saveRDS(params_by_age, file = "../saves/params_by_age.rds")

model_params_by_age <- params_by_age%>%
  select(value) %>%
  map_df(bind_rows)%>%
  group_by(age_bin)%>%
  summarise(so_map = estimate_mode(speakerOptimality),
            diff_priorReg_map = estimate_mode(diff_priorReg),
            same_priorReg_map = estimate_mode(same_priorReg))

@@

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}

var ex3Data = dataFromR.ex3data
var modelParams = dataFromR.modelParams

    display(JSON.stringify(ex3Data))

var predict_by_age = function(age_bin){

    //display(JSON.stringify(age_bin))

    var speakerOptimality = age_bin.so_map

    var diff_priorReg = age_bin.diff_priorReg_map

    var same_priorReg =age_bin.same_priorReg_map

  return extend(
      {rsaPredictions_diff_cong,
      rsaPredictions_diff_incong,
      rsaPredictions_same_cong,
      rsaPredictions_same_incong,
      age_bin
    })

 }

var predictions_by_age = map(predict_by_age, modelParams)

predictions_by_age

'
```

#model_pred <- readRDS("../saves/kids_model_predictions.rds")

model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, modelPredWebppl , sep='\n'),
  data = list(ex3data = ex3_data, modelParams = model_params_by_age),
  data_var = "dataFromR",
  model_var = "predict_by_age",
  chains = 1,
  cores = 4,
  inference_opts = list(method = "MCMC", samples = 10, burn = 1, verbose = T)
)

@@

  // global parameters

@@

@@

    return [ 
    [ age_bin + "_diff_cong",  noisyRsaPredictions_diff_cong],
    [ age_bin + "_diff_incong",  noisyRsaPredictions_diff_incong],
    [ age_bin + "_same_cong",  noisyRsaPredictions_same_cong],
    [ age_bin + "_same_incong",  noisyRsaPredictions_same_incong],
]

@@

return _.fromPairs(predictions_by_age.concat([
["so_slope", so_slope],
["so_int", so_int],
["noise", noise]
]))

}
'
```
```{r noisy combined model predictions}

@@

# Sanity checks

@@

## importing data
ex3_data <- read_csv(file="../data/kids_ex3_data.csv")

### 2019-02-08 stats/kids/kids_pref_2.Rmd
update kids model preictions

```{r setup, include= FALSE}
knitr::opts_chunk$set(echo = FALSE)
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)

## importing data

# d_ipad <- read_csv(file="kids_pref.data.csv")
# 
# d_hand_code <- read.csv(file="kids_pref_handcoding.data.csv", sep = ";")%>%
#   mutate(age_num = as.numeric(as.character(age_num)))
# 
# d <- bind_rows(d_ipad,d_hand_code)%>%
#   mutate(trial_type = ifelse(trial == "train", "train", "test"))%>%
#   filter(subid != "200718_8_pref")
# 
# 
# # check if someone needs to be excluded because wrong in training
# # no one
# d %>%
#   filter(trial_type == "train") %>%
#   group_by(subid)%>%
#   summarise(correct = mean(correct)) %>%
#   filter(correct == 0)

d <- read_csv(file="data/kids_pref_data.csv")

@@

# Subjects

There was one additional 4yo in the data set, whom I excluded to get the pre-registered sample size of 45 total. For the modeling we might want to include her as that gives us a tiny bit more precision.

## Overall

All training trials were answered correctly, suggesting that children generally understood the procedure. Sex is not included in the data set but the sex ration is fairly balances across age bins (47 to 60% girls).

```{r sanity checks, echo = FALSE,warning = FALSE, message = FALSE}

## sanity checks
d %>%
  group_by(trial_type) %>%
  summarise(n = length(unique(subid)), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 2)

@@

## By age bin

Just to check if we got the planned sample size of 15 per bin.

```{r kids per age group,echo = FALSE}

d %>%
  group_by(age)%>%
  summarise(n = length(unique(subid)),
            age_num = mean(age_num))%>%
  knitr::kable(digits = 2)
```

@@

## Plot across age bins

Medium overall effect of speaker change.

```{r plot across age}
p3 <- d %>%
  filter(trial_type == "test") %>%
  group_by(change, subid) %>%
  summarise(correct = mean(correct))

p4 <- p3 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p3, aes(x = change, y = correct, col = change, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p4, aes(x = change, y = mean, col = change, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

@@

```{r plot by age}
p1 <- d %>%
  filter(trial_type == "test") %>%
  group_by(change,age, subid) %>%
  summarise(correct = mean(correct))

@@

ggplot() +
  geom_jitter(data = p1, aes(x = change, y = correct, col = change, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = change, y = mean, col = change, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~age)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

>>>>

```{r setup, include= FALSE}
knitr::opts_chunk$set(echo = FALSE)
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)

## importing data
files <- dir("~/Work/MCC/git-mcc/kids_pref_data_2/")

 raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_pref_data_2/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,45,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

length(unique(raw_data$subid))

d_raw <-raw_data %>%
  mutate(test_date = as.Date(test_date),
         experiment = "mcc_kids_pref_2")

log <- read_excel("../../../MCC-subject_log.xlsx", 1)%>%
  filter(Condition == "pref2")%>%
  select(subid,experimenter,keep_drop,sex,dob)

length(unique(log$subid))

sum(log$keep_drop=="drop")

d <- left_join(d_raw,log, by = "subid")%>%
  filter(keep_drop == "keep")%>%
  mutate(dob = as.numeric(dob),
         dob = as.Date(dob,origin = "1899-12-30"),
         age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))

length(unique(d$subid))

@@

# Sample size per age bin

```{r kids per age group,echo = F }

d %>%
  group_by(subage)%>%
  summarise(n = length(unique(subid)))%>%
  knitr::kable(digits = 2)

@@

@@

@@

```{r plot by age}
p1 <- d %>%
  filter(trial != "train") %>%
  group_by(speakerChange,subage, subid) %>%
  summarise(correct = mean(correct))

@@

ggplot() +
  geom_jitter(data = p1, aes(x = speakerChange, y = correct, col = speakerChange, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = speakerChange, y = mean, col = speakerChange, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```
